---
title: "Is an automatic or manual transmission better for MPG?"
author:
output:
  pdf_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 4
geometry: margin=0.6in
---

```{r echo=FALSE}
library(ggplot2)
library(plyr)
library(lattice)
library(MASS)
library(car)
library(caret)
library(grid)
```

### Executive summary  
To answer the stated question exactly we should collect MPG values 
from all existing brands (total population). With the help of regression 
analysis and a sample, we can answer this question approximately, while quantifying uncertainty
in our prediction. A simple regression of MPG against transmission type shows that 
**manual transmission better for MPG by 7.25 miles per gallon on average**.
Adding more variables increases model's ability to predict MPG of a car.  

### Exploratory analysis  

Let's explore data visually by plotting MPG separately for two groups of cars: with automatic
and manual transmission (see Appendix, Figure 1: MPG vs transmission type).
The cursory visual analysis does suggest that there is a difference in MPG due to
transmission type. Let's validate this conjecture by regression analysis.  

### Regression analysis 1.  

Let's perform an Ordinary Least Squares (OLS) regression of MPG ("mpg") on transmission type ("am").
It should be noted that although MPG is a continuous variable, transmission type is
categorical one that can take on two states: "automatic" or "manual". To perform
OLS I am treating "am" as dummy variable that can take on two values: "0" for "automatic" and "1" for manual (factorization explicit or implicit also possible). Summary of resulting regression model:

```{r echo=FALSE}
ls1 <- lm(mpg~am, data=mtcars)
summary(ls1)
```

### Regression 1: Coefficient interpretation  

- Intercept is mean MPG for automatic transmission (set am=0 and get average
MPG for automatic transmission).
- `am` coefficient is incremental increase in MPG due to switch to manual transmission. In other words, **this is MPG difference between automatic and manual transmissions**. As before, if you set `am=1`, you will get mean MPG for cars with manual transmission. You can check these statements by comparing coefficients to calculated means:
```{r}
ddply(mtcars, "am", function(x) mean(x$mpg))
```
Both coefficients -- `(Intercept)` and `am` -- are statistically significant as 
evidenced by low p-values, i.e. mean MPG for automatic transmission significantly different
from 0; and mean MPG for manual transmission significantly different from mean for automatic transmission (in other words, both have predictive values).  

### Regression 1: Quantifying model uncertainty and regression diagnostics  

Predicted average MPG for automatic and manual transmission with 95% Confidence intervals are:
```{r}
cbind(am = c(0,1), rbind(predict(ls1, newdata=data.frame(am=0), interval = "confidence"),
                         predict(ls1, newdata=data.frame(am=1), interval = "confidence")))
```



Normality of residuals distribution is not a problem here, as evidenced by 
Q-Q plot (see Appendix, Figure 2: Residuals Q-Q plot for Regression 1). The two
real problems are:  
- low explanatory ability of the model (adjusted $R^2 = 0.3385$)  
- patterns in residuals distributions. See Appendix, Figure 3: Patterns in residual distribution by transmission type and additional variable, as an example of one pattern: heavier cars tend to have
negative residuals (MPG overprediction).

### Regression analysis 2: Adding more explanatory variables  

There are two paths to adding more explanatory variables:  
- **Manual**. First add to regression model all predictors that visually have linear
relationships to MPG (see Appendix, Plot 3, for visualization of
relationships between MPG and predictors), and then remove by hand
insignificant ones, starting with least significant, unless only
significant predictors are left in the model.  
- **Automatic**. Let `stepAIC()` function from `MASS` package choose the "best" regression
model for you ( as the name of the function suggests, this is done via minimizing Akaike information criterion).
As it turns out in this particular case, both approaches end up in identical model:

```{r echo=FALSE}
lsTotal <- lm(mpg~., data=mtcars)
ls2 <- stepAIC(lsTotal, trace=F)
summary(ls2)
```

### Regression 2: interpretation, diagnostics, and comparison to Regression 1  

- All predictors, except intercept, are significant
- 1 sec increase in `qsec` ("1/4 mile time") will increase MPG by 1.23 while all other variables kept constant; 1000lb increase in weight will decrease MPG by 3.92; and manual transmission cars will have 2.94 Miles per gallon higher MPG on average over those with automatic 
transmission.
- Regression 2 has higher predictive ability as measured by adjusted $R^2$: 0.8336 vs 0.3385
- We are 95% confident that residuals are normal (see Appendix, Figure 5: Residuals Q-Q plot for Regression 2)
- Plot of residuals against regression components still reveals some heterescedasticity (see Appendix, Figure 6: Component + Residual Plots for Regression 2)

# Appendix
```{r fig.width=4, fig.height=3.8, echo=FALSE, fig.cap="MPG vs transmission type", }
data(mtcars)
ggplot(data=mtcars, aes(x=factor(am), y=mpg)) +
        geom_boxplot(aes(group=factor(am), color=factor(am))) +
        labs(x="transmission type", y = "MPG, miles/(US) gallon") +
        theme(axis.title.x = element_text(size=9),
              axis.title.y = element_text(size=9)) +
        scale_color_discrete(name="transmission type",
                             labels=c("0 - Automatic", "1 - Manual"))
```



```{r echo=FALSE, fig.cap="Residuals Q-Q plot for Regression 1"}
qqPlot(ls1$resid)

```

```{r echo=FALSE, fig.cap="Patterns in residual distribution by transmission type and additional variable"}
ggplot(data=NULL, aes(x=factor(mtcars$am), y=resid(ls1))) +
        geom_violin(aes(color=factor(mtcars$am))) +
        geom_point(aes(fill=mtcars$wt), shape=21, size=2) +
      labs(x="transmission type", y = "MPG residuals") +
        theme(axis.title.x = element_text(size=9),
              axis.title.y = element_text(size=9)) +
        scale_color_discrete(name="transmission type",
                             labels=c("0 - Automatic", "1 - Manual")) +
        scale_fill_continuous(name="Car weight")
```


```{r echo=FALSE, fig.cap = "Feature plot: MPG vs all predictors", fig.width=5, fig.height=5}
featurePlot(x=mtcars[,!(names(mtcars) == "mpg")], y=mtcars[,"mpg"])
```


```{r echo=FALSE, fig.cap="Residuals Q-Q plot for Regression 2"}
qqPlot(ls2$resid)
```

```{r echo=FALSE, fig.cap="Component + Residual Plots for Regression 2", warning=FALSE, fig.height=5, fig.width=5}
crPlots(ls2)
```
